---
title: "PSY6422_Data_Management_&_Visualization"
author: "Monika Sobieraj"
date: "24.05.2021"
output: 
  html_document:  default
---
***
 ![](pokemon_vector.svg) 

***

# Gotta Plot 'Em All!


## Research Question

This visualization attempts to find our how Pokemon's strength changed throughout generations (1-6) and if not-legendary Pokemon can defeat legendary, and what are they?

## Pokemon Information

> generations:

**I Generation** - (1996-1999) New Pokemon - 151

**II Generation** - (1999-2001) New Pokemon - 100

**III Generation** - (2002-2005) New Pokemon - 135

**IV Generation** - (2006-2010) New Pokemon - 107

**V Generation** - (2010-2013) New Pokemon - 156

**VI Generation** - (2013-2016) New Pokemon - 72

> legendary:

Legendary Pokemon are incredibly rare and often very powerful Pokemon, generally featured prominently in the legends and myths of the Pokemon world.

*information source:* https://bulbapedia.bulbagarden.net/wiki/Legendary_Pok%C3%A9mon *and* https://bulbapedia.bulbagarden.net/wiki/Generation .

## Data Origins

The data consists of 721 Pokemon and specific for each of them attributes such as HP, Attack, Type.1 etc.

Pokemon data that was used for this project and can be found at: [pokemon data](https://www.kaggle.com/abcsds/pokemon)

```{r include=FALSE}

if(!require("tidyverse"))install.packages("tidyverse")
if(!require("dplyr"))install.packages("dplyr")
if(!require("tidyr"))install.packages("tidyr") 
if(!require("ggplot2"))install.packages("ggplot2")
if(!require("tibble"))install.packages("tibble")
if(!require("plotly"))install.packages("plotly") 
if(!require("wesanderson"))install.packages("wesanderson") 
if(!require("gganimate"))install.packages("gganimate") 
if(!require("htmlwidgets"))install.packages("htmlwidgets") 

library(tidyverse)
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(plotly)
library(wesanderson)
library(gganimate)
library(htmlwidgets)
```

```{r}
data <- read.csv("data/pokemon_data.csv") #load data

#there are 13 variables which are:
names(data)
```
## Data Preparation

```{r}
#tidying and removing all except variables we need
df <- data %>% tibble::as_tibble() %>% 
   distinct(X., .keep_all = TRUE) %>% 
  select(Name, Type.1, Total, Generation, Legendary) %>%
   dplyr::rename(type = Type.1,
         name = Name,
         total = Total,
         generation = Generation,
         legendary = Legendary) %>%
  arrange(type) %>%
  print() 

#tidying, changing type
df$legendary[df$legendary=="True"] <- "Legendary" 
df$legendary[df$legendary=="False"] <- "Not Legendary"

head(df,10)
```
## Visualisation 1: Rainbow Pokemon - All


```{r}
g1 <- df %>%
  mutate(text=paste("Name: ",(df$name), "\nType: ",(df$type), "\nTotal: ", (df$total), sep="")) %>% 
  ggplot(df,mapping = aes(x = total, y=type, col=type, text=text)) +
  geom_point(position = "jitter")+
  labs( x="Total strength",
        y= NULL,
        title="Types of Pokemon")+
  theme_minimal()+ 
  theme(axis.text.y=element_blank(),
        plot.title = element_text(size=20))+ 
  guides(col=guide_legend("Types of Pokemon"))

p1 <- ggplotly(g1, tooltip="text") 
p1

#save output
htmlwidgets::saveWidget(p1, file="graphs/rainbow_pokemon.html")
```



### Visualisation 2: Legendary Pokemon
```{r}
names(wes_palettes)

g2 <- df %>%
mutate(text=paste("Name: ",(df$name), "\nType: ",(df$type), "\nTotal: ", 
                  (df$total), "\nGen: ", (df$generation), "\n", 
                  (df$legendary),sep="")) %>%
ggplot(df,mapping = aes(x = generation, y=total, col=legendary, text=text)) +
  geom_point(position = "jitter")+
  labs( x=NULL,
        y= "Total strength",
        title="Legendary Pokemon")+
  theme_minimal()+
  theme(axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(size=20),
        panel.grid.major.x = element_blank())+
  guides(col=guide_legend("Types of Pokemon"))+
  scale_colour_manual(values=wes_palette("Moonrise3"))

p2 <- ggplotly(g2, tooltip="text") 
p2

#save output
htmlwidgets::saveWidget(p2, file="graphs/legendary_pokemon.html")
```
### Visualisation 3 - Interactive Pokemon Types
```{r}
g3 <- df %>%
  mutate(text=paste("Name: ",(df$name), "\nType: ",(df$type), "\nTotal: ", 
                    (df$total), "\nGen: ", (df$generation), "\n", 
                    (df$legendary),sep="")) %>%
  ggplot(df,mapping = aes(x = generation, y=total, col=type, text=text)) +
  geom_point(position = "jitter")+
  labs( x=NULL,
        y= "Total strength",
        title="All Pokemon")+
  theme_minimal()+
  theme(axis.text.x=element_blank(),
          plot.title = element_text(size=20),
          panel.grid.major.y = element_line("#e6f2ff"),
          panel.grid.major.x = element_blank(), 
          axis.line = element_line(colour = "#003166"))+
  guides(col=guide_legend("Types of Pokemon"))+
  scale_color_manual(values = c("Bug"="tan4","Dark"="black","Dragon"="chartreuse4","Electric"="darkgoldenrod1","Fairy"="hotpink","Fighting"="darksalmon","Fire"="firebrick1","Flying"="lightskyblue","Ghost"="darkblue","Grass"="yellowgreen", "Ground"="peru","Ice"="cadetblue1","Normal"="pink","Poison"="maroon","Psychic"="darkviolet","Rock"="wheat3","Steel"="gray","Water"="steelblue1"))

p3 <- ggplotly(g3, tooltip="text") 
p3

#save output
htmlwidgets::saveWidget(p3, file="graphs/all_types_interactive.html")
```



### Visualisation 4: Animated Pokmon All
```{r}
g4 <- ggplot(df,mapping = aes(x = generation, y=total, col=type)) +
  geom_point(position = "jitter", size=2)+
  labs(title="Now showing {closest_state} Pokemon",
       subtitle = 'Frame {frame} of {nframes}',
       x=NULL,
       y= "Total strength")+
  theme_minimal()+
  theme(axis.text.x=element_blank(),
        axis.title.y = element_text(size=15),
        plot.title = element_text(size=20),
        panel.grid.major.y = element_line("#e6f2ff"),
        axis.line = element_line(colour = "#003166"),
        legend.position = "none")+ #removes legend
  transition_states(type,
                    transition_length = 20,
                    state_length = 10,
                    wrap = FALSE)+
  enter_fade()+ 
  exit_fade()+
  scale_color_manual(values = c("Bug"="tan4","Dark"="black","Dragon"="chartreuse4","Electric"="darkgoldenrod1","Fairy"="hotpink","Fighting"="darksalmon","Fire"="firebrick1","Flying"="lightskyblue","Ghost"="darkblue","Grass"="yellowgreen", "Ground"="peru","Ice"="cadetblue1","Normal"="pink","Poison"="maroon","Psychic"="darkviolet","Rock"="wheat3","Steel"="gray","Water"="steelblue1"))

animate(g4 + ease_aes(), width=600,height=600) 

#save output
anim_save(file="graphs/all_types_animated.gif") 
```
## Summary Discussion

Caveat #1: Type.2 of Popkemon were not look at, which might be interesting to do so in the future.

Caveat #2:  2 new generations (VII[2016-2019] and VIII [2019-2022]) were not included in the data and there are 177 new Pokemon.

Caveat #3:  Comparison of each attributes of Total Score such as HP and Attack might be worth looking at
